Direct Spectroscopic Detection: A Cheap Method to Detect and Characterize Binary Systems

### 2015-10-08 paper/ms.tex
Updated section 5.

We use Equation \ref{eqn:gyro} to estimate the expected rotation period for a companion star of given temperature and age as follows: First, we convert from temperature to convective timescale ($\tau$) by interpolating Table 1 in \citet{Barnes2010a}. Next we sample an appropriate PDF for the age of the binary system; if the primary star was analyzed in \citet{David2015}, we use their posterior age PDFs. Otherwise, we use a uniform PDF from the ZAMS age of the primary star to its main sequence lifetime (typically 10-200 Myr for our sample). Following the discussion in \cite{Barnes2010b}, we uniformly sample initial rotation periods from 0.2 - 5 days for all stars. We estimate the current rotation period for each pair of age and initial rotation period samples using Equation \ref{eqn:gyro} to build up a PDF of current rotation periods. We transform the period distribution into a PDF for $v\sin{i}$ using the Main Sequence radius of a star of the given temperature, obtained by interpolating Table 1 of \cite{Barnes2010a}, and a uniform sampling of inclinations ($\sin{i}$). We show a typical vsini distribution in Figure \ref{fig:vsini}, which peaks near $\sim 5-10$ $\rm km \ s^{-1}$ and has a long tail extending to $\sim 40-50 \ \rm km s^{-1}$.

@@

By combining the sensitivity calculations described above with the $v\sin{i}$ samples, we marginalize over the expected rotation periods of the secondary stars to get simpler curves of detection rate vs companion star temperature. We show the median and approximate range of the marginalized detection rate in Figure \ref{fig:sensitivity}. The direct spectral detection method can reliably detect companions as cool as 3700 K in most cases, although the primary star spectral type plays a dominant role in setting the coolest detectable companion.

Companions with $T \gtrsim 6250$, the canonical limit at which the convective zone is too small to transfer angular momentum to the stellar wind and spin down the star \citep{Pinsonneault2001}, may have rotational velocities comparable to that of the primary star. In that case, estimating the primary star spectrum with a gaussian filter may remove much or all of the companion spectrum. Since these are the types of stars with less extreme flux- and mass-ratios, they are easier to detect with more conventional methods. However, this shortcoming could be overcome by using model spectra for the primary star as in \citet{Kolbl2015}. In this work, we have optimized the method for finding cool companions.

@@

We now use the direct spectroscopic detection method to measure the temperatures of several known binary systems (Table \ref{tab:known}). We cross-correlate the spectra against the full grid of model spectra enumerated in Section \ref{sec:method}, and find the temperature using Equation \ref{eqn:tmeas}. We then convert the measured temperature to posterior probability distributions (pdfs) for the true companion temperature using the MCMC chains developed in Section \ref{subsec:systematics}. For stars with more than one detection, we multiply the pdfs from each detection. Finally, we calculate the final companion temperature and confidence interval from the integral of the pdf:

>>>>

We use Equation \ref{eqn:gyro} to estimate the expected rotation period for a companion star of given temperature and age as follows: First, we convert from temperature to convective timescale ($\tau$) by interpolating Table 1 in \citet{Barnes2010a}. Next we sample an appropriate probability density function (PDF) for the age of the binary system; if the primary star was analyzed in \citet{David2015}, we use their posterior age PDFs. Otherwise, we use a uniform PDF from the Zero Age Main Sequence (ZAMS) age of the primary star to its main sequence lifetime (typically 10-200 Myr for our sample). Following the discussion in \cite{Barnes2010b}, we uniformly sample initial rotation periods from 0.2 - 5 days for all stars. We estimate the current rotation period for each pair of age and initial rotation period samples using Equation \ref{eqn:gyro} to build up a PDF of current rotation periods. We transform the period distribution into a PDF for $v\sin{i}$ using the main sequence radius of a star of the given temperature, obtained by interpolating Table 1 of \cite{Barnes2010a}, and a uniform sampling of inclinations ($\sin{i}$). Figure \ref{fig:vsini} shows a typical $v\sin{i}$ distribution, which peaks near $\sim 5-10$ $\rm km \ s^{-1}$ and has a long tail extending to $\sim 40-50 \ \rm km s^{-1}$.

@@

By combining the sensitivity calculations described above with the $v\sin{i}$ samples, we marginalize over the expected rotation periods of the secondary stars to get simpler curves of detection rate as a function of companion star temperature. We show the median and approximate range of the marginalized detection rate in Figure \ref{fig:sensitivity}. The direct spectral detection method can reliably detect companions as cool as 3700 K in most cases, although the primary star spectral type plays a dominant role in setting the coolest detectable companion.

Companions with $T \gtrsim 6250$ K, the canonical limit at which the convective zone is too small to transfer angular momentum to the stellar wind and spin down the star \citep{Pinsonneault2001}, may have rotational velocities comparable to that of the primary star. In that case, estimating the primary star spectrum with a gaussian filter may remove much or all of the companion spectrum. Since these are the types of stars with less extreme flux- and mass-ratios, they are easier to detect with more conventional methods. However, this shortcoming could be overcome by using model spectra for the primary star as in \citet{Kolbl2015}. In this work, we have optimized the method for finding cool companions.

@@

We now use the direct spectroscopic detection method to measure the temperatures of several known binary systems (Table \ref{tab:known}). We cross-correlate the spectra against the full grid of model spectra enumerated in Section \ref{sec:method}, and find the temperature using Equation \ref{eqn:tmeas}. We then convert the measured temperature to PDFs for the true companion temperature using the MCMC chains developed in Section \ref{subsec:systematics}. For stars with more than one detection, we multiply the PDFs from each detection. Finally, we calculate the final companion temperature and confidence interval from the integral of the PDF:
